---
title: "ECON 121 FA23 Problem Set 2"
author: "Your name"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

\newpage

## Question 2

Verbal answer only, no code.

\newpage

## Question 3

Code: Load packages and dataset, modify sample, generate variables, summarize data.

Verbal: Interpret the summary statistics.

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, modify sample, generate variables here.

```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.

```

\newpage

## Question 4

Code: Estimate regression.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 4 code here

```

\newpage

## Question 5

Code: Estimate regression.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 5 code here

```

\newpage

## Question 6

Code: Assess statistical significance.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 6 code here

```

\newpage

## Question 7

Code: Estimate regressions and assess significance of differences between them.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 7 code here

```

\newpage

## Question 8

Code: Estimate regression, use delta method.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

```

\newpage

## Question 9

Code: Load dataset, compute unweighted and weighted means.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 9 code here

```

\newpage

## Question 10

Code: Modify sample, estimate unweighted and weighted regressions.

Verbal: Assess unweighted versus weighted results.

```{r echo=TRUE}
# All question 10 code here

```

\newpage

## Question 11

Verbal answer only, no code.

\newpage

## Question 12

Verbal answer only, no code.

\newpage

## Question 13

Code: Estimate regression.

Verbal: Explain choice of covariates, interpret results.

```{r echo=TRUE}
# All question 13 code here

```

