---
title: "ECON 121 FA23 Problem Set 3"
author: "Your name"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

\newpage

## Question 2

Code: Load packages and dataset,generate variables, summarize data.

Verbal: Interpret the summary statistics.

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, generate variables here.

```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.

```

\newpage

## Question 3

Code: Draw graph with two line plots.

Verbal: Interpret.

```{r echo=TRUE}
# All question 3 code here.

```

\newpage

## Question 4

Code: Draw bar graphs.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 4 code here

```

\newpage

## Question 5

Code: Estimate regressions.

Verbal: Interpret and compare.

```{r echo=TRUE}
# All question 5 code here

```

\newpage

## Question 6

Code: Use the results from the mortality logit model to compare the two groups.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 6 code here

```

\newpage

## Question 7

Verbal: Assess causality.

\newpage

## Question 8

Code: Assess how much health behavior can explain the mortality logit results.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

```

