---
title: "ECON 121 FA23 Final Exam"
author: "Your name"
output: pdf_document
---

```{r setup, include=FALSE}
# set the code chunk format
knitr::opts_chunk$set(echo = TRUE)
# clear the environment
rm(list=ls())
# load all packages from the course
library(tidyverse)
library(fixest)
library(car)
library(mfx)
# load the dataset
#load(url("https://github.com/tvogl/econ121/raw/main/data/migration.Rdata"))
```

## Question 1

```{r}
# Your code here
```

\newpage

## Question 2

Your verbal answer here

\newpage

## Question 3

```{r}
# Your code here
```

\newpage

## Question 4

```{r}
# Your code here
```

\newpage

## Question 5

Your verbal answer here

\newpage

## Question 6

```{r}
# Your code here
```

\newpage

## Question 7

```{r}
# Your code here
```

\newpage

## Question 8

```{r}
# Your code here
```

\newpage

## Question 9

Your verbal answer here

\newpage

## Question 10

```{r}
# Your code here
```

\newpage

## Question 11

```{r}
# Your code here
```

\newpage

## Question 12

Your verbal answer here

\newpage

## Question 13

```{r}
# Your code here
```

\newpage

## Question 14

Your verbal answer here

\newpage

## Question 15

Your verbal answer here

\newpage

## Question 16

Your verbal answer here

\newpage

## Question 17

```{r}
# Your code here
```

\newpage

## Question 18

Your verbal answer here

\newpage

## Question 19

```{r}
# Your code here
```

\newpage

## Question 20

Your verbal answer here







