---
title: "ECON 121 FA23 Problem Set 1"
author: "Robert Tso"                                                                                                                                #:)#
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.
Robert Tso - A13829791

\newpage

## Question 2

Code: Load packages and dataset, summarize the data.

Verbal: Interpret the summary statistics.

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset here.

library(readr)

#tinytex::install_tinytex()

install.packages('tidyverse', repos = "http://cran.us.r-project.org")
library(tidyverse)

install.packages('plyr', repos = "http://cran.us.r-project.org")
library(dplyr)

#ssa_names <- read_csv("D:/Documents/Class/Econ 121/econ121/data/ssa_names.csv")
ssa_names <- read_csv("C:/Users/hp/Documents/GitHub/econ121/data/ssa_names.csv")
# View(ssa_names)

```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.

summary(ssa_names)

#There are 1,718,855 name entries, ranging from years of 1940 to 2022. 
#A minimum of 5 occurrences for any name and maximum of 99,693 occurrences.

```

\newpage

## Question 3

Code: Find the all time most common boy and girl names.

Verbal: Interpret your results.

```{r echo=TRUE}

#aggregate by name
total_names <- ssa_names%>%
  group_by(sex,name) %>%
  summarize(total_frequency = sum(frequency))%>%
  ungroup()



total_names

#girls
total_girl_names <- total_names%>%
  filter(sex=="F") #filter by sex for female

#filter total girl names by the most frequent name
total_girl_names %>% filter(total_frequency==max(total_frequency)) 


#store the frequency for girls
most_common_girl <- max(total_girl_names$total_frequency) 



#boys
total_boy_names <- total_names%>%
  #filter by sex for male
  filter(sex=="M") 

#filter total boy names by the most frequent name
total_boy_names %>% filter(total_frequency==max(total_frequency)) 

#store the frequency for boy
most_common_boy <- max(total_boy_names$total_frequency)

#add the total frequencies
print(most_common_girl+most_common_boy)
#Total of 6,261,170 babies of the 2 most popular names.

# Mary was the most popular name for girls at 1,997,025 
# and Michael for boys at 4,264,145 A total of 6,261,170 babies 
# since 1940 were born under these 2 popular names.
```

\newpage

## Question 4

Code: Plot time trends in the number of unique names by sex

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 4 code here

unique_girls_by_year <- ssa_names %>%
  filter(sex=="F")%>%
  group_by(year) %>%
  #count the amount of distinct names each year for girls
  summarise(unique_name_count = n_distinct(name)) 

#unique_girls_by_year

unique_boys_by_year <- ssa_names %>%
  filter(sex=="M")%>%
  group_by(year) %>%
  #count the amount of distinct names each year for boys
  summarise(unique_name_count = n_distinct(name))

#unique_boys_by_year

unique_names_plot <- ggplot() +
  geom_line(data = unique_boys_by_year, aes(x = year, y = unique_name_count, color = "Boys")) +
  geom_line(data = unique_girls_by_year, aes(x = year, y = unique_name_count, color = "Girls")) +
  
  labs(x = "Year", y = "Number of Unique Names", title = "Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Boys" = "blue", "Girls" = "salmon")) +
  theme_minimal()

unique_names_plot

# Throughout the years, girls have had more unique names than boys, 
# and they trend in similar ways. The number of unique names for both boys 
# has remained relatively flat compared to girls until 1960-1970s. 
# Both sexes peaked in uniqueness in 2008, where boys returned to a flat rate, 
# while girls started declining.

```

\newpage

## Question 5

Code: Plot time trends in the number of unique names relative to the number of 
babies by sex.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 5 code here

ratio_girl_babies <- ssa_names %>%
  filter(sex=="F")%>%
  group_by(year) %>%
  summarise(ratio = n_distinct(name)/sum(frequency))
#ratio_girl_babies

ratio_boy_babies <- ssa_names %>%
  filter(sex=="M")%>%
  group_by(year) %>%
  summarise(ratio = n_distinct(name)/sum(frequency))
#ratio_boy_babies


ratio_names_plot <- ggplot() +
  geom_line(data = ratio_boy_babies, aes(x = year, y = ratio, color = "Boys")) +
  geom_line(data = ratio_girl_babies, aes(x = year, y = ratio, color = "Girls")) +
  
  labs(x = "Year", y = "Ratio", title = "Ratio of Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Boys" = "blue", "Girls" = "salmon")) +
  theme_minimal()

ratio_names_plot

#The ratio of unique names increases over time, with slight declines 
# during certain time periods, such as the mid-1970s to mid-1980s, and after 
#2010, for girls a steady decline while for boys only a short decline. 
# The patterns display a decreasing of conformity among babies. 

```

\newpage

## Question 6

Verbal: Which names and why?

Apple because it was the first word I could think of that starts with "A". 
Sam because this is a gender neutral name, short for Samuel or Samantha. 

\newpage

## Question 7

Code: Find all-time frequency of your names and compare with all-time most
popular names.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 7 code here

#apple
Apple_names <- ssa_names%>%
  filter(name=="Apple")
Apple_names_girls <- Apple_names%>%
  filter(sex=="F")
Apple_names_boys <- Apple_names%>%
  filter(sex=="M")

Apple_names_girls
Apple_names_boys

Apple_names_plot_log <- ggplot() +
  geom_line(data = unique_boys_by_year, aes(x = year, y = unique_name_count, color = "Boys")) +
  geom_line(data = unique_girls_by_year, aes(x = year, y = unique_name_count, color = "Girls")) +
  geom_line(data = Apple_names_boys, aes(x = year, y = frequency, color = "Apple Boys")) +
  geom_line(data = Apple_names_girls, aes(x = year, y = frequency, color = "Apple Girls"))+
  
  scale_y_continuous(trans='log10')+           #Scaled the graph into Log 
  labs(x = "Year", y = "Number of Unique Names", title = "Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Boys" = "blue", "Girls" = "salmon","Apple Boys"="purple","Apple Girls"="red")) +
  theme_minimal()

#sam
Sam_names <- ssa_names%>%
  filter(name=="Sam")
Sam_names_girls <- Sam_names%>%
  filter(sex=="F")
Sam_names_boys <- Sam_names%>%
  filter(sex=="M")

Sam_names_girls
Sam_names_boys


Sam_names_plot_log <- ggplot() +
  geom_line(data = unique_boys_by_year, aes(x = year, y = unique_name_count, color = "Boys")) +
  geom_line(data = unique_girls_by_year, aes(x = year, y = unique_name_count, color = "Girls")) +
  geom_line(data = Sam_names_boys, aes(x = year, y = frequency, color = "Sam Boys")) +
  geom_line(data = Sam_names_girls, aes(x = year, y = frequency, color = "Sam Girls"))+
  
  scale_y_continuous(trans='log10')+           #Scaled the graph into Log 
  labs(x = "Year", y = "Number of Unique Names", title = "Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Boys" = "blue", "Girls" = "salmon","Sam Boys"="green","Sam Girls"="orange")) +
  theme_minimal()

Sam_names_plot <- ggplot() +
  geom_line(data = unique_boys_by_year, aes(x = year, y = unique_name_count, color = "Boys")) +
  geom_line(data = unique_girls_by_year, aes(x = year, y = unique_name_count, color = "Girls")) +
  geom_line(data = Sam_names_boys, aes(x = year, y = frequency, color = "Sam Boys")) +
  geom_line(data = Sam_names_girls, aes(x = year, y = frequency, color = "Sam Girls"))+
  
  labs(x = "Year", y = "Number of Unique Names", title = "Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Boys" = "blue", "Girls" = "salmon","Sam Boys"="green","Sam Girls"="orange")) +
  theme_minimal()

Apple_names_plot_log
#I had not checked beforehand how many boys were named Apple, 
#so I was surprised to see no line representing boys named Apple

Sam_names_plot_log
#I was surprised to see how erratic, but constant Sam's name was for girls, 
#meanwhile for boys, Sam was a slow decline.

Sam_names_plot
#I checked again without the Log, and see that Sam is near non-existent 
#among girls, while Sam was still relatively popular before the 1960s for boys. 

```

\newpage

## Question 8

Code: Graph the annual frequencies of your chosen name(s) over time.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

picked_names_plot <- ggplot() +
  geom_line(data = Apple_names_boys, aes(x = year, y = frequency, color = "Apple Boys")) +
  geom_line(data = Apple_names_girls, aes(x = year, y = frequency, color = "Apple Girls"))+
  geom_line(data = Sam_names_boys, aes(x = year, y = frequency, color = "Sam Boys")) +
  geom_line(data = Sam_names_girls, aes(x = year, y = frequency, color = "Sam Girls"))+
  
  scale_y_continuous(trans='log10')+           #Scaled the graph into Log 
  labs(x = "Year", y = "Number of Unique Names", title = "Unique Names by Year (Boys vs. Girls)") +
  scale_color_manual(values = c("Apple Boys"="purple","Apple Girls"="red","Sam Boys"="green","Sam Girls"="orange")) +
  theme_minimal()

picked_names_plot
#Sam and Apple are similar in popularity among girls, as they are very 
# unconventional names, and Apple only started appearing after the 1970s, 
# predictably more popular among  hippie/non-conformist parents. Sam as a name 
# for boys sees a steady decline after 1960, which is the same time as the near
# exponential increase of non-conformity among all baby names mentioned in
# question 5.

```