---
title: "ECON 121 FA23 Problem Set 2"
author: "Robert Tso"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.
Robert Tso - A13829791
Tomas Lopez - 
Stephanie Nguyen - 

\newpage

## Question 2

Beta_0 is the intercept of the formula
Beta_1 is the coefficient for the level of education

\newpage

## Question 3

Code: Load packages and dataset, modify sample, generate variables, summarize data.

Verbal: Interpret the summary statistics.

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, modify sample, generate variables here.
library(tidyverse)
```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.
load("D:/Documents/Class/Econ 121/econ121/data/cps18.Rdata")
summary(cps18)
#drop rows with fewer than 50 weeks, worked fewer than 35 hours in a typical week, or has 0 dollars in annual earnings
cps <- cps18 %>% 
  filter(
  wkswork>=50 &
  hrs_per_wk>=35 &
  incwage > 0
  )
  
#add log hourly wage variable
 cps <- cps%>%
  mutate(
    annual_hours=wkswork*hrs_per_wk,
    hr_wage=incwage/annual_hours,
    log_hr_wage=log(hr_wage)
  )
 
#add race/ethnicity dummy variables
Black <- ifelse(cps$race == 'black', 1, 0)
Native <- ifelse(cps$race == 'native', 1, 0)
Asian <- ifelse(cps$race == 'asian/pacific', 1, 0)
Other <- ifelse(cps$race == 'multiple/other', 1, 0)

#add the dummy's to the cps dataframe
cps <- cps%>%
  mutate(
    Black=Black,
    Native=Native,
    Asian=Asian,
    Other=Other
  )

#add up the education columns with weights of the total years spent.
#since these are dummy variables, only one column for each row should be true
cps <- cps %>%
  mutate(education = 
        ed_lt_hs*8 + #less than high school is assuming middle school completion
        ed_some_hs*10 + #some high school is 2 years or half for total 10
        ed_hs_degree*12 + #completing high school is total 12 years
        ed_some_col*14 + #some college is 2 years or half for total 14
        ed_ba_degree*16 + #BA is 16 years total       
        ed_post_degree*18 #post degree is 2 years for 18 total
        )
#generate exper = age - education - 5
cps <- cps %>%
  mutate(exper = age - education - 5)

#generate exper2 = exper^2
cps <- cps %>%
  mutate(exper2 = exper^2)

view(cps)
summary(cps)
#The data's median is a 43 year old white male with 12 years of education.
#The oldest person 64 years old and given a maximum experience of 51, 
#only has 8 years of education
#The max annual hours worked is over 4 times the median, 
#and considering that the max and min weeks worked is only 2 apart
```

\newpage

## Question 4

Code: Estimate regression.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 4 code here

```

\newpage

## Question 5

Code: Estimate regression.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 5 code here

```

\newpage

## Question 6

Code: Assess statistical significance.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 6 code here

```

\newpage

## Question 7

Code: Estimate regressions and assess significance of differences between them.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 7 code here

```

\newpage

## Question 8

Code: Estimate regression, use delta method.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

```

\newpage

## Question 9

Code: Load dataset, compute unweighted and weighted means.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 9 code here

```

\newpage

## Question 10

Code: Modify sample, estimate unweighted and weighted regressions.

Verbal: Assess unweighted versus weighted results.

```{r echo=TRUE}
# All question 10 code here

```

\newpage

## Question 11

Verbal answer only, no code.

\newpage

## Question 12

Verbal answer only, no code.

\newpage

## Question 13

Code: Estimate regression.

Verbal: Explain choice of covariates, interpret results.

```{r echo=TRUE}
# All question 13 code here

```

