---
title: "ECON 121 FA23 Problem Set 2"
author: "Robert Tso"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

Robert Tso - A13829791

Tomas Lopez - A16215540

Stephanie Nguyen - A16798775

Akash Juwadi - A16372772

\newpage

## Question 2

Beta_0 is the intercept of the formula. Beta_1 is the coefficient that indicates the hourly wage return for each increase in level of education. Experience squared makes total experience a quadratic and looks to see if "excess" experience is a detriment to hourly wage, such as being too old or not having enough education.

\newpage

## Question 3

Code: Load packages and dataset, modify sample, generate variables, summarize data.

Verbal: Interpret the summary statistics.

The data's median is a 43 year old white male with 12 years of education, there is not much skew in the age and education. The max annual hours worked is over 4 times the median, and considering that the max and min weeks worked is only 2 apart

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, modify sample, generate variables here.
library(tidyverse)
library(fixest)
library(car)


```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.
load("D:/Documents/Class/Econ 121/econ121/data/cps18.Rdata")
summary(cps18)

#drop rows with fewer than 50 weeks, worked fewer than 35 hours in a typical week, 
#or has 0 dollars in annual earnings
cps <- cps18 %>% 
  filter(
  wkswork>=50 &
  hrs_per_wk>=35 &
  incwage > 0
  )
  
#calculate annual hours and hourly wage
annual_hours <- cps$wkswork*cps$hrs_per_wk
hr_wage <- cps$incwage/annual_hours

#add log hourly wage variable
 cps <- cps%>%
  mutate(
    log_hr_wage=log(hr_wage)
  )
 
#add race/ethnicity dummy variables
Black <- ifelse(cps$race == 'black', 1, 0)
Native <- ifelse(cps$race == 'native', 1, 0)
Asian <- ifelse(cps$race == 'asian/pacific', 1, 0)
Other <- ifelse(cps$race == 'multiple/other', 1, 0)

#add the dummy's to the cps dataframe
cps <- cps%>%
  mutate(
    Black=Black,
    Native=Native,
    Asian=Asian,
    Other=Other
  )

#add up the education columns with weights of the total years spent.
#since these are dummy variables, only one column for each row should be true
cps <- cps %>%
  mutate(education = 
        ed_lt_hs*8 + #less than high school is assuming middle school completion
        ed_some_hs*10 + #some high school is 2 years or half for total 10
        ed_hs_degree*12 + #completing high school is total 12 years
        ed_some_col*14 + #some college is 2 years or half for total 14
        ed_ba_degree*16 + #BA is 16 years total       
        ed_post_degree*20 #post degree is 2 years for 18 total
        )

#generate exper = age - education - 5
cps <- cps %>%
  mutate(exper = age - education - 5)

#generate exper2 = exper^2
cps <- cps %>%
  mutate(exper2 = exper^2)

#view(cps)
summary(cps)

```

\newpage

## Question 4

Code: Estimate regression.

Verbal: Interpret your results.

{log_hr_wage = 1.190526 + 0.106590*education + 0.027118*exper - 0.000397*exper2 }

Intercept only accounts for 1.190526 of the hourly wage, regardless of age, education or experience.Experience seems to have a relatively lower effect compared to education, but a reminder is that education is subtracted from age to get experience.Each year of Education has a 10.7% increase of hourly wage.

```{r echo=TRUE}
# All question 4 code here
# Define the Mincerian wage model

mince_model <- feols(log_hr_wage ~ education + exper + exper2, data = cps)

# Print the summary of the regression model
summary(mince_model)

```

\newpage

## Question 5

Code: Estimate regression.

Verbal: Interpret your results.

log_hr_wage = 1.029361 + 0.109797*education + 0.025147*exper - 0.000357*exper2	
            - 0.158205*black - 0.109256*Native + 0.028961*Asian - 0.055523*Other
            + 0.273182*male
            
Every year of Education is now 10.97% increase in wage, which is higher. 



```{r echo=TRUE}
# All question 5 code here

ext_mince_model <- feols(log_hr_wage ~ education + exper + exper2  
                        + Black + Native + Asian + Other
                        + male, data = cps)

# Print the summary of the regression model
summary(ext_mince_model)


```

\newpage

## Question 6

Code: Assess statistical significance.

Verbal: Interpret your results.

The Z-value is 40.904, which is really high, therefore it is 
statistically significant


```{r echo=TRUE}
# All question 6 code here

deltaMethod(ext_mince_model, "male - Black", rhs=0)

```

\newpage

## Question 7

Code: Estimate regressions and assess significance of differences between them.

Verbal: Interpret your results.

The T-value for the two coefficients of male and female education is 2.300352, 
which is statistically significant.

```{r echo=TRUE}
# All question 7 code here

gender_mince_model <- feols(log_hr_wage ~ education + exper + exper2  
                        + Black + Native + Asian + Other,
                        data = cps, split = ~male)

gender_mince_model

#testing coefficients for education
(0.112598 - 0.107783)/sqrt((0.001547^2) + (0.001410^2))
```

\newpage

## Question 8

Code: Estimate regression, use delta method.

Verbal: Interpret your results.

The difference between the coefficients for male education and female education is 
0.004815, which is exactly the coefficient of malexedu 0.004816.

Education, Experience, Exper2, and Black are the only ratios that are 
statistical significant from 1.


```{r echo=TRUE}
# All question 8 code here

#building the interaction terms
cps_interactions <- mutate(cps, malexedu = male*education,
                                malexexper = male*exper,
                                malexexper2 = male*exper2,
                                blackmale = male*Black,
                                nativemale = male*Native,
                                asianmale = male*Asian,
                                othermale = male*Other)

#building extended mincerian model with interaction terms
ext_gender_mince_model <- feols(log_hr_wage ~ education + exper + exper2  
                        + Black + Native + Asian + Other + male
                        + malexedu + malexexper + malexexper2 
                        + blackmale + nativemale 
                        + asianmale + othermale,
                        data = cps_interactions)

summary(ext_gender_mince_model)
summary(gender_mince_model)

#extended mince model for 
female_mince_model <- feols(log_hr_wage ~ education + exper + exper2  
                        + Black + Native + Asian + Other,
                        data = cps,  subset = ~male==0)

male_mince_model <- feols(log_hr_wage ~ education + exper + exper2  
                        + Black + Native + Asian + Other,
                        data = cps,  subset = ~male==1)

#deltaMethod test for each variable between male and female
deltaMethod(ext_gender_mince_model,"education/(education + malexedu)",RHS=1)
deltaMethod(ext_gender_mince_model,"exper/(exper + malexexper)",RHS=1)
deltaMethod(ext_gender_mince_model,"exper2/(exper2 + malexexper2)",RHS=1)

deltaMethod(ext_gender_mince_model,"Black/(Black + blackmale)",RHS=1)

deltaMethod(ext_gender_mince_model,"Native/(Native + nativemale)",RHS=1)
deltaMethod(ext_gender_mince_model,"Asian/(Asian + asianmale)",RHS=1)
deltaMethod(ext_gender_mince_model,"Other/(Other + othermale)",RHS=1)


#divide the coefficients directly to get a ratio, but no staistical significance.
coefficients(female_mince_model)/coefficients(male_mince_model)



```

\newpage

## Question 9

Code: Load dataset, compute unweighted and weighted means.

Verbal: Interpret your results.

The summary gives us the mean of black at 0.2502 and mean of hisp at 0.1578. When weighed black mean is 0.1386967 and hisp mean is 0.063047. When accounting for sampling weights, the estimates bias would be reduced.

```{r echo=TRUE}
# All question 9 code here

load("D:/Documents/Class/Econ 121/econ121/data/nlsy79.Rdata")
summary(nlsy79)
#view(nlsy79)

weighted.mean(nlsy79$black,w=nlsy79$perweight)
weighted.mean(nlsy79$hisp,w=nlsy79$perweight)


```

\newpage

## Question 10

Code: Modify sample, estimate unweighted and weighted regressions.

Verbal: Assess unweighted versus weighted results.

When we account for the sampling weights, the magnitudes of all coefficients 
and intercepts EXCEPT for male, decrease, meanwhile male increases from 0.327012 to 0.390222.
The swing of coefficients decreases from unweighted to weighted are marginal compared 
to the swing of coefficient increase for male.

We decided to stick with the unweighted dataset as the relatively drastic 
increase of male coefficient takes away from the importance of the other variables




```{r echo=TRUE}
# All question 10 code here

#drop rows with fewer than 50 weeks times 35 hours = 1750 annual hours worked, 
#or has 0 dollars in annual earnings
nlsy <- nlsy79 %>% 
  filter(
  hours18>=1750 &
  laborinc18 > 0
  )
  
#calculate hourly wage
hr_wage_79 <- nlsy$laborinc18/nlsy$hours18

#add log hourly wage variable
 nlsy <- nlsy%>%
  mutate(
    log_hr_wage=log(hr_wage_79)
  )

#generate exper = age - education - 5
nlsy <- nlsy %>%
  mutate(exper = age79 - educ - 5)

#generate exper2 = exper^2
nlsy <- nlsy %>%
  mutate(exper2 = exper^2)
 
 
summary(nlsy)

unweigh_nlsy = feols(log_hr_wage ~ educ + exper + exper2 + male + black + hisp, data = nlsy)
weighed_nlsy = feols(log_hr_wage ~ educ + exper + exper2 + male + black + hisp, data = nlsy, weights = ~perweight)

unweigh_nlsy
weighed_nlsy

```

\newpage

## Question 11

Verbal answer only, no code.

The coefficients for education and experience are different, CPS measures more impact from experience while nsly has more impact with education. Differences in coefficients should be expected as there are different explanatory variables that are not accounted in the mincerian equation from each respective dataset. For example nlsy parental data, sources of magazines, library, newspapers, which can supplement overall experience. There is also significant differences between the demographics of the datasets. The average age of CPS is 43 while for nsly it is 17, demonstrating a bias towards older and more established careers as someone who is 30+ may have more career opportunities than someone just graduating high school.

```{r echo=TRUE}
#comparing coefficients directly
coefficients(ext_mince_model)
coefficients(unweigh_nlsy)

#comparing data summaries
summary(cps)
summary(nlsy)

```

\newpage

## Question 12

Verbal answer only, no code.

It is not a guarantee that education is a causal effect on hourly wages. As we have columns for childhood environment that affect a child's propensity for education, such as access to a library and educated parents. Any of these variables may possibly be the true causal effect that affect the education coefficient. 



\newpage

## Question 13

Code: Estimate regression.

Verbal: Explain choice of covariates, interpret results.

At first glance AFQT may seem important as it measures job qualifications, however AFQT contains some overlaps with an IQ test, so it may just be a redundant approximation for education. We can observe that the education coefficient drops when accounting for afqt. Similarly the value of education from mom and dad have a positive impact at a marginal cost of the education coefficient, as parents that value education would probably impart that value onto their child, or at the very least force them to go to school regardless. When accounting for all 3 of these variables, education coefficient drops even more, indicating an importance to control AFQT as it is responsible for the largest swing.

```{r echo=TRUE}
# All question 13 code here

AFQT_nlsy = feols(log_hr_wage ~ educ + exper + exper2 
                              + male + black + hisp
                              + afqt81,
                                data = nlsy)

dad_nlsy = feols(log_hr_wage ~ educ + exper + exper2 
                              + male + black + hisp
                              + educ_dad,
                                data = nlsy)

mom_nlsy = feols(log_hr_wage ~ educ + exper + exper2 
                              + male + black + hisp
                              + educ_mom,
                                data = nlsy)

all_nsly = feols(log_hr_wage ~ educ + exper + exper2 
                              + male + black + hisp
                              + afqt81 + educ_dad + educ_mom,
                                data = nlsy)

coefficients(unweigh_nlsy)
coefficients(AFQT_nlsy)
coefficients(dad_nlsy)
coefficients(mom_nlsy)
coefficients(all_nsly)


```

