---
title: "ECON 121 FA23 Problem Set 3"
author: "Robert Tso"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

Robert Tso - A13829791
Tomas Lopez - A16798775
Stephanie Nguyen - A16215540
Akash Juwadi - A16372772

\newpage

## Question 2

Code: Load packages and dataset, generate variables, summarize data.

Verbal: Interpret the summary statistics.

The summary statistics show a seemingly near population pool of survey answers, except there is a slight weight towards female answers, as the population split of male/female in the US is closer to 49/51% as opposed to the 43/57% split in the dataset. The dummy variables we generated for fair and poor health show that 16% of the survey judge themselves in this classification. It is worth nothing that this percent is near and above the range of percentage answers for diabetic(12%) and alcohol use(11%).

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, generate variables here.
library(tidyverse)
library(fixest)

view(nhis2010)
load(url("https://github.com/tvogl/econ121/raw/main/data/nhis2010.Rdata"))
#load("C:/Users/hp/Documents/GitHub/econ121/data/nhis2010.Rdata")

# drop observations with health missing/NA.
nhis2010 <- nhis2010 %>% drop_na(health)

# generate a variable that equals one if fair or poor health, zero otherwise.
table(nhis2010$health)
nhis2010$health_dummy <- ifelse(nhis2010$health %in% c("Fair", "Poor"), 1, 0)

#the sum of Fair and Poor should be the same as 1
table(nhis2010$health_dummy)


```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.
summary(nhis2010)



```

\newpage

## Question 3

Code: Draw graph with two line plots.

Verbal: Interpret.

Risk of death for both categories go up with age, however the greatest difference of mortality between the self-reported health groups is more pronounced between ages 40 and 80, with a clear observation of lower risk of death among those with self-reported good-to-excellent health. In the beginning and end of the data, both groups have very similar mortality rates.

```{r echo=TRUE}
# All question 3 code here.

# Compute mortality rates by age for both groups
mortality_data <- nhis2010 %>%
  drop_na(age, mort)%>%
  group_by(health_dummy, age) %>%
  summarise(mortality_rate = sum(mort) / n())

# Create separate line plots for the two groups
ggplot(mortality_data, aes(x = age, y = mortality_rate, color = factor(health_dummy))) +
  geom_line() +
  labs(
    x = "Age",
    y = "Mortality Rate",
    title = "Mortality Rate by Age for Different Health Groups",
    color = "Health Group"
  )



```

\newpage

## Question 4

Code: Draw bar graphs.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 4 code here

```

\newpage

## Question 5

Code: Estimate regressions.

Verbal: Interpret and compare.

```{r echo=TRUE}
# All question 5 code here

```

\newpage

## Question 6

Code: Use the results from the mortality logit model to compare the two groups.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 6 code here

```

\newpage

## Question 7

Verbal: Assess causality.

\newpage

## Question 8

Code: Assess how much health behavior can explain the mortality logit results.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

```

