---
title: "ECON 121 FA23 Problem Set 3"
author: "Robert Tso"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

Robert Tso - A13829791

Tomas Lopez - A16798775

Stephanie Nguyen - A16215540

Akash Juwadi - A16372772

\newpage

## Question 2

Code: Load packages and dataset, generate variables, summarize data.

Verbal: Interpret the summary statistics.

The summary statistics show a seemingly near population pool of survey answers, except there is a slight weight towards female answers, as the population split of male/female in the US is closer to 49/51% as opposed to the 43/57% split in the dataset. The dummy variables we generated for fair and poor health show that 16% of the survey judge themselves in this classification. It is worth nothing that this percent is near and above the range of percentage answers for diabetic(12%) and alcohol use(11%).

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, generate variables here.
#install.packages("mfx")
library(mfx)

#install.packages("betareg")
library(betareg)

library(tidyverse)
library(fixest)
library(car)

#load(url("https://github.com/tvogl/econ121/raw/main/data/nhis2010.Rdata"))
load("D:/Documents/Class/Econ 121/econ121/data/nhis2010.Rdata")
#view(nhis2010)

# drop observations with health missing/NA.
nhis2010 <- nhis2010 %>% drop_na(health)

# generate a variable that equals one if fair or poor health, zero otherwise.
table(nhis2010$health)
nhis2010$health_dummy <- ifelse(nhis2010$health %in% c("Fair", "Poor"), 1, 0)

#the sum of Fair and Poor should be the same as 1
table(nhis2010$health_dummy)


```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.
summary(nhis2010)



```

\newpage

## Question 3

Code: Draw graph with two line plots.

Verbal: Interpret.

Risk of death for both categories go up with age, however the greatest difference of mortality between the self-reported health groups is more pronounced between ages 40 and 80, with a clear observation of lower risk of death among those with self-reported good-to-excellent health. In the beginning and end of the data, both groups have very similar mortality rates.

```{r echo=TRUE}
# All question 3 code here.

# Compute mortality rates by age for both groups
mortality_data <- nhis2010 %>%
  drop_na(age, mort)%>%
  group_by(health_dummy, age) %>%
  summarise(mortality_rate = mean(mort))

# Create separate line plots for the two groups
ggplot(mortality_data, aes(x = age, y = mortality_rate, color = factor(health_dummy))) +
  geom_line() +
  labs(
    x = "Age",
    y = "Mortality Rate",
    title = "Mortality Rate by Age for Different Health Groups",
    color = "Health Group"
  )



```

\newpage

## Question 4

Code: Draw bar graphs.

Verbal: Interpret your results.

In the case of family income, as income increases, rates of mortality and poor health decreases. Similarly as education level increases, rates of mortality and poor health decreases. We are not certain of the cross interaction of education and income on either mortality or health, but it would not be surprising if the interactive case was true. Looking at the odds ratios, all the incomes higher than $35,000 have lower odds of mortality and better reported health, meanwhile all education lower than college have higher odds. In regards to race differences, Asians and Hispanics have lower mortality rates compared to Whites, and Blacks and Other have a higher rate to claim poor/fair health relative to Whites.

```{r echo=TRUE}
# All question 4 code here

# Create table for fair/poor health and mortality by family income
graph_a <- nhis2010 %>%
  drop_na(incfam,mort) %>%
  group_by(incfam) %>%
  summarise(mean_fair_poor_health = mean(health_dummy),
            mean_mortality = mean(mort))

# Create bar plots for family income
fam_health <- ggplot(graph_a, aes(x = incfam)) +
  geom_bar(aes(y = mean_fair_poor_health), stat = "identity", fill = "blue", position = "dodge") +
  labs(x = "Family Income", y = "Mean Value", title = "Rates of Fair/Poor Health by Family Income")

fam_mort <- ggplot(graph_a, aes(x = incfam)) +
  geom_bar(aes(y = mean_mortality), stat = "identity", fill = "red", position = "dodge") +
  labs(x = "Family Income", y = "Mean Value", title = "Rates of Mortality by Family Income")

# Categorize years of education into five categories
nhis2010 <- nhis2010 %>%
  drop_na(edyrs)%>%
  mutate(education_category = case_when(
    edyrs < 12 ~ "Less than High School",
    edyrs == 12 ~ "High School Completion",
    edyrs >= 13 & edyrs <= 15 ~ "Some College",
    edyrs == 16 ~ "College Completion",
    edyrs > 16 ~ "Post-graduate Study"
  ))

# Create table for fair/poor health and mortality by education category
graph_b <- nhis2010 %>%
  drop_na(mort)%>%
  group_by(education_category) %>%
  summarise(mean_fair_poor_health = mean(health_dummy),
            mean_mortality = mean(mort))

# Create bar plots for education
edu_health <- ggplot(graph_b, aes(x = education_category)) +
  geom_bar(aes(y = mean_fair_poor_health), stat = "identity", fill = "blue", position = "dodge") +
  labs(x = "Education Level", y = "Mean Value", title = "Rates of Fair/Poor Health by Education Level") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))  # Rotate x-axis labels

edu_mort <- ggplot(graph_b, aes(x = education_category)) +
  geom_bar(aes(y = mean_mortality), stat = "identity", fill = "red", position = "dodge") +
  labs(x = "Education Level", y = "Mean Value", title = "Rates of Mortality by Education Level") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))  # Rotate x-axis labels


fam_health
fam_mort
edu_health
edu_mort

```

\newpage

## Question 5

Code: Estimate regressions.

Verbal: Interpret and compare.

Based on the summary statistics OLS does not look like a good predictive model because it has negative values on the lower end of the model. It also has a dampened maximum value at close to 50%, compared to the 66-70%+ of logit and probit. 

```{r echo=TRUE}
# All question 5 code here

#dropping empty rows
nhis2010 <- nhis2010 %>%
  drop_na(mort,health_dummy,incfam,age,education_category,black,hisp,asian,other) 

#generating dummy variables for education categories
LessHS <- ifelse(nhis2010$education_category == "Less than High School", 1, 0)
HSGrad <- ifelse(nhis2010$education_category == "High School Completion", 1, 0)
SomeCol <- ifelse(nhis2010$education_category == "Some College", 1, 0)
ColGrad <- ifelse(nhis2010$education_category == "College Completion", 1, 0)
PostGrad <- ifelse(nhis2010$education_category == "Post-graduate Study", 1, 0)

#generating dummy variables for income categories
Low <- ifelse(nhis2010$incfam == "$0 - $34,999", 1, 0)
LowMed <- ifelse(nhis2010$incfam == "$35,000 - $49,999", 1, 0)
Med <- ifelse(nhis2010$incfam == "$50,000 - $74,999", 1, 0)
MedHigh <- ifelse(nhis2010$incfam == "$75,000 - $99,999", 1, 0)
High <- ifelse(nhis2010$incfam == "$100,000 and over", 1, 0)


#add all the dummy variables to nhis2010
nhis2010 <- nhis2010 %>%
  mutate(
    LessHS=LessHS,
    HsGrad=HSGrad,
    SomeCol=SomeCol,
    ColGrad=ColGrad,
    PostGrad=PostGrad,
    Low=Low,
    LowMed=LowMed,
    Med=Med,
    MedHigh=MedHigh,
    High=High
  )  

#view(nhis2010)

ols_model_pf <- feols(health_dummy ~  age +
                      Low + LowMed + Med + MedHigh + High +
                      LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                      black + hisp + asian + other, 
                      data = nhis2010, 
                      vcov = 'hetero')
nhis2010$ols_predict_pf <- predict(ols_model_pf, nhis2010, type="response")


ols_model_mort <- feols(mort ~  age +
                        Low + LowMed + Med + MedHigh + High +
                        LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                        black + hisp + asian + other, 
                        data = nhis2010, 
                        vcov = 'hetero')
nhis2010$ols_predict_mort <- predict(ols_model_mort, nhis2010, type="response")

probit_model_pf <- feglm(health_dummy ~  age +
                        Low + LowMed + Med + MedHigh + High +
                        LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                        black + hisp + asian + other, 
                        data = nhis2010, 
                        vcov = 'hetero',
                        family = 'probit')
nhis2010$probit_predict_pf <- predict(probit_model_pf, nhis2010, type="response")

probit_model_mort <- feglm(mort ~  age +
                          Low + LowMed + Med + MedHigh + High +
                          LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                          black + hisp + asian + other, 
                          data = nhis2010, 
                          vcov = 'hetero',
                          family = 'probit')
nhis2010$probit_predict_mort <- predict(probit_model_mort, nhis2010, type="response")


logit_model_pf <- feglm(health_dummy ~  age +
                        Low + LowMed + Med + MedHigh + High +
                        LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                        black + hisp + asian + other, 
                        data = nhis2010, 
                        vcov = 'hetero',
                        family = 'logit')
nhis2010$logit_predict_pf <- predict(logit_model_pf, nhis2010, type="response")


logit_model_mort <- feglm(mort ~  age +
                          Low + LowMed + Med + MedHigh + High +
                          LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                          black + hisp + asian + other, 
                          data = nhis2010, 
                          vcov = 'hetero',
                          family = 'logit')
nhis2010$logit_predict_mort <- predict(logit_model_mort, nhis2010, type="response")

# ols_model_pf
# probit_model_pf
# logit_model_pf

#summary statistics of health
nhis2010 %>% 
  select(ols_predict_pf,logit_predict_pf,probit_predict_pf) %>%
  summary()

#summary statistics of mortality
nhis2010 %>% 
  select(ols_predict_mort,logit_predict_mort,probit_predict_mort) %>%
  summary()

#marginal effect of IVs for poor and fair health using logit
logitmfx(health_dummy ~  incfam + age + education_category + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
#marginal effect of IVs for poor and fair health using probit
probitmfx(health_dummy ~  incfam + age + education_category + black + hisp + asian + other,  
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)

#marginal effect of IVs for mortality using logit
logitmfx(mort ~  incfam + age + education_category + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
#marginal effect of IVs for poor and fair healt usin probit
 probitmfx(mort ~  incfam + age + education_category + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
 
 
 #odds ratio of logit mortality
logitor(mort ~incfam+ age + education_category + black + hisp + asian + other, 
        data = nhis2010, 
        robust = TRUE)

#odds ratio of logit poor or fair health
logitor(health_dummy ~incfam+ age + education_category + black + hisp + asian + other, 
        data = nhis2010, 
        robust = TRUE)


```

\newpage

## Question 6

Code: Use the results from the mortality logit model to compare the two groups.

Verbal: Interpret your results.

Given these scenarios, Group A has a greater mortality rate, which makes sense since education and income are stronger predictors of mortality than race. We should include interaction terms because Asian adults with a low education and low income and Black adults with college graduate education and over $100k incomes since both collectively represent less than 1% of the population of the data and the basic logit models would not be good predictors of such small data subsets.

```{r echo=TRUE}
# All question 6 code here

#looking at the percent of people in the survey that are of Group A or B
nhis2010 %>% 
  filter(asian*LessHS*Low==1) %>%
  count()/count(nhis2010)  
#0.375% are Group A

nhis2010 %>% 
  filter(black*ColGrad*High==1) %>%
  count()/count(nhis2010)
#0.247% are Group B


logit_model_mort

deltaMethod(logit_model_mort,"(asian + LessHS + Low) - (black + ColGrad)",rhs=0)

#manually
#Group A is an asian adult with less than 12 years or education and family income less the 35k
GroupA <- (-0.580831612  #asian coef
          +0.755474   #less than $35k
          +0.362800)	 #less than 12 years edu

# Group B: Black adults with 16 years of education and family incomes over $100k
GroupB <- (-0.003324	 #black coef
          +0.211033 # 16 years of edu
          +0) #family income over 100k, 0 due to collinearity

GroupA-GroupB



```

\newpage

## Question 7

Verbal: Assess causality.

No, there may be an omitted variable that predicts both income and health, such as motivation to work harder in health and in career. This motivation variable may be impacted by socioeconomic or purely random and normally distributed through the population.

\newpage

## Question 8

Code: Assess how much health behavior can explain the mortality logit results.

Verbal: Interpret your results.

When measured separately, being Uninsured has a coefficient of -0.074457 while smoking has a coefficient of 0.544864. Measured together uninsured is -0.101424 and smoking is 0.544725, showing that smoking has a relatively large increase in mortality, while being uninsured has a lower, but negative effect. Using the odds ratio we see that smoking has  1.7241335, which means smoking has a 72% higher likelihood of mortality versus those that do not smoke. Similarly being uninsured has a 10% higher likelihood of mortality. 

```{r echo=TRUE}
# All question 8 code here

#model for uninsured on mortality
logit_model_uninsur <- feglm(mort ~  age + uninsured + 
                            Low + LowMed + Med + MedHigh + High +
                            LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                            black + hisp + asian + other, 
                            data = nhis2010, 
                            vcov = 'hetero',
                            family = 'logit')
#odds ratio of logit uninsured
logitor(logit_model_uninsur, 
        data = nhis2010, 
        robust = TRUE)

logit_model_uninsur

#model for smoking on mortality
logit_model_smoke <- feglm(mort ~  age + smokev + 
                            Low + LowMed + Med + MedHigh + High +
                            LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                            black + hisp + asian + other, 
                            data = nhis2010, 
                            vcov = 'hetero',
                            family = 'logit')
#odds ratio of logit drinking
logitor(logit_model_smoke, 
        data = nhis2010, 
        robust = TRUE)

logit_model_smoke

#model for uninsured and smoking on mortality
logit_model_unsur_smoke <- feglm(mort ~  age + uninsured + smokev + 
                            Low + LowMed + Med + MedHigh + High +
                            LessHS + HsGrad + SomeCol + ColGrad + PostGrad +
                            black + hisp + asian + other, 
                            data = nhis2010, 
                            vcov = 'hetero',
                            family = 'logit')
logitor(logit_model_unsur_smoke, 
        data = nhis2010, 
        robust = TRUE)

logit_model_unsur_smoke


```

