---
title: "ECON 121 FA23 Problem Set 3"
author: "Tomas Lopez"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

\newpage

## Question 2

Code: Load packages and dataset,generate variables, summarize data.

Verbal: The mean age of respondednts to the NHS is about 51, with the median
being 49. This is a weighted sample where weights are given to those with less
common attributes of race marital status etc. The sample is made up of primarily 
white indviduals, at 57%.  It is also 43% Male and and 57% Female. 13% of the 
sample died by 2019.

```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset, generate variables here.
install.packages('mfx')
install.packages('betareg')
library(tidyverse)
library(fixest)
library(mfx)
library(betareg)
library(car)


load(url("https://github.com/tvogl/econ121/raw/main/data/nhis2010.Rdata"))

summary(nhis2010)

nhis2010 <- nhis2010 %>% drop_na(health, mort)
#dropping na values in health column

#dummy var for fair and poor health condition
table(nhis2010$health)
nhis2010 <- nhis2010 %>%
  mutate(poor_fair = case_when(
    health == "Poor" ~ 1,   # if health status is poor
    health == "Fair" ~ 1,   # if health status is fair 
    TRUE ~ 0               # Default value if neither condition is met
  ))
table(nhis2010$poor_fair)





```

```{r echo=TRUE}
# The PDF will show the code AND output here.
summary(nhis2010)

```

\newpage

## Question 3

Code: Draw graph with two line plots.

Verbal: The risk of death increases across the board by age, regardless of what
an self individuals perceived health is.  However those who do percieve their 
health worse are more likely to die than those who are in percieve themselves to
be in better health, although the gap does level off at when looking aat ages 
past 80.

```{r echo=TRUE}
# All question 3 code here.

mortality_data <- nhis2010 %>%
  drop_na(age, mort)%>%
  group_by(poor_fair, age) %>%
  summarise(mortality_rate = mean(mort))

# Create separate line plots for the two groups
ggplot(mortality_data, aes(x = age, y = mortality_rate, color = factor(poor_fair))) +
  geom_line() +
  labs(
    x = "Age",
    y = "Mortality Rate",
    title = "Mortality Rate by Age for Different Health Groups",
    color = "Health Group"
  )


```
\newpage

## Question 4

Code: Draw bar graphs.

Verbal:In graph (a) the percentage of poor or fair health self reported 
individuals decreases with income, that is family income and self reported 
health are negatively correlated.  In graph (b), we see a negative correlation 
between education level and individuals who self report poor or fair health.

```{r echo=TRUE}
# All question 4 code here

#dropping na vals of income
income_poor_fair_no_na <- nhis2010 %>% drop_na(incfam)

#grouping by income and then taking a percentage of those poor_fair in each
#income bracket

income_poor_fair <- income_poor_fair_no_na %>%
  group_by(incfam) %>%
  summarise(percent_fair_poor = sum(poor_fair) / n())

print(income_poor_fair)



# Create the bar graph using ggplot
ggplot(income_poor_fair, aes(x = incfam, y = percent_fair_poor)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(
    x = "Income Level",
    y = "Percentage of Fair or Poor Health",
    title = "Percentage of Fair or Poor Health by Income Level"
  )


#creating new column based on education 
nhis2010 <- nhis2010 %>%
  drop_na(edyrs) %>%
  mutate(educational_attainment = case_when(
    edyrs < 12 ~ "Less than High School Completion",
    edyrs == 12 ~ "High School Completion",
    edyrs >= 13 & edyrs <= 15 ~ "Some College",
    edyrs == 16 ~ "College Completion",
    edyrs > 16 ~ "Post-graduate Study",
  ))
view(nhis2010)
edyrs_no_na <- nhis2010 %>% drop_na(edyrs)

#grouping by income and then taking a percentage of those poor_fair in each
#income bracket

edyrs_poor_fair <- edyrs_no_na %>%
  group_by(educational_attainment) %>%
  summarise(percent_edyrs_pf = sum(poor_fair) / n())

print(edyrs_poor_fair)

ggplot(edyrs_poor_fair, aes(x = educational_attainment, y = percent_edyrs_pf)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(
    x = "Education",
    y = "Percentage of Fair or Poor Health",
    title = "Percentage of Fair or Poor Health by Income Level"
  )+ theme(axis.text.x = element_text(size = 6))

```

\newpage

## Question 5

Code: Estimate regressions.

Verbal: We expect post graduate to be the least likely group to have poor or fair health,
as well as lowest mortaility odds. Comparing across the three modesl OLS is clearly
the worst, it reports negative values for mortality, which not interpretable.
Looking at the odds ratio, all income levels higher than 35k there are lower odds of mortality and poor or 
fair self reported health, when looking at education levels lower than college completion, we
see higher odds of mortality and poor or fair self reported health. Amongst, 
Hispanics and Asians show lower odds of mortality relative white people, when
looking at Blacks and Others they show higher odds of poor and fair health 
relative to white people.

```{r echo=TRUE}
# All question 5 code here
#ols linear model

nhis2010 <- nhis2010 %>%
  drop_na(incfam,age,educational_attainment,black,hisp,asian,other)

ols_model_pf <- feols(poor_fair ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero')
nhis2010$ols_predict_pf <- predict(ols_model_pf, nhis2010, type="response")

ols_model_mort <- feols(mort ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero')
nhis2010$ols_predict_mort <- predict(ols_model_mort, nhis2010, type="response")

probit_model_pf <- feglm(poor_fair ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero',
                   family = 'probit')

nhis2010$probit_predict_pf <- predict(probit_model_pf, nhis2010, type="response")

probit_model_mort <- feglm(mort ~ incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero',
                   family = 'probit')
nhis2010$probit_predict_mort <- predict(probit_model_mort, nhis2010, type="response")



logit_model_pf <- feglm(poor_fair ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero',
                   family = 'logit')
nhis2010$logit_predict_pf <- predict(logit_model_pf, nhis2010, type="response")


logit_model_mort <- feglm(mort ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                   vcov = 'hetero',
                   family = 'logit')
nhis2010$logit_predict_mort <- predict(logit_model_mort, nhis2010, type="response")


ols_model_pf
probit_model_pf
logit_model_pf


summary(nhis2010$ols_predict_pf)
summary(nhis2010$logit_predict_pf)
summary(nhis2010$probit_predict_pf)
summary(nhis2010$poor_fair)

summary(nhis2010$ols_predict_mort)
summary(nhis2010$logit_predict_mort)
summary(nhis2010$probit_predict_mort)
summary(nhis2010$mort)



#marginal effect of IVs for poor and fair health using logit
logitmfx(poor_fair ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
#marginal effect of IVs for poor and fair healt usin probit
probitmfx(poor_fair ~  incfam+ age + educational_attainment + black + hisp + asian + other,  
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)

#marginal effect of IVs for mortality using logit
logitmfx(mort ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
#marginal effect of IVs for poor and fair healt usin probit
 probitmfx(mort ~  incfam+ age + educational_attainment + black + hisp + asian + other, 
                   data = nhis2010, 
                  atmean = TRUE, 
                  robust = TRUE)
 
 
 #odds ratio of logit mortality
logitor(mort ~incfam+ age + educational_attainment + black + hisp + asian + other, 
        data = nhis2010, 
        robust = TRUE)

#odds ratio of logit poor or fair health
logitor(poor_fair ~incfam+ age + educational_attainment + black + hisp + asian + other, 
        data = nhis2010, 
        robust = TRUE)


view(nhis2010)


```

\newpage

## Question 6

Code: Use the results from the mortality logit model to compare the two groups.

Verbal: Group A is worse off in terms of mortality, we expect this since the key
predictors from question 5 are education and income, NOT race.  To improve we 
could include interaction terms, 

```{r echo=TRUE}
# All question 6 code here

-0.02801763 
0.02532388
coef(logit_model_mort)


#Group A is an asian adult with less than 12 years or education and fam income less the 35k
print(-0.580831612  #asian coef
      +0.417197392 #less than 12 years edu
      +0)

# Group B: Black adults with 16 years of education and family incomes over $100k
print( -0.003323531 #black coef
      +0.054397497 # 16 or more years edu
      -0.755473560 ) #family income over 100k





```

\newpage

## Question 7

Verbal: Assess causality.

\newpage

## Question 8

Code: Assess how much health behavior can explain the mortality logit results.

Verbal: Interpret your results.

```{r echo=TRUE}
# All question 8 code here

```

