---
title: "ECON 121 FA23 Problem Set 4"
author: "Robert Tso"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # sets the code chunk format
rm(list = ls()) # clears the environment
```

## Question 1

Verbal: list group members.

Robert Tso - A13829791

Tomas Lopez - A16798775

Stephanie Nguyen - A16215540

Akash Juwadi - A16372772

ChatGPT

\newpage

## Question 2

Code: Load packages and dataset, summarize data.

Verbal: Interpret the summary statistics.

Notable mean differences between the headstart participating children are sibdiff of 0.350750326, showing that children who are in headstart have a third likelihood to have siblings also in headstart. There are also 25% more black children but 4.7% less hispanic children. The largest impact is comp_score_5to6	with a mean difference of -5.560310055	and median difference of-6.0, showing that headstart children have lower test scores in this bracket. Logged income or lninc_0to3	with mean of -0.421336775	and median of -.5775361, showing that headstart children have 0.5	less logged income.


```{r message=FALSE, warning=FALSE}
# The PDF will show the code you write here but not the output.
# Load packages and dataset here.

rm(list=ls())
library(tidyverse)
library(fixest)

load(url("https://github.com/tvogl/econ121/raw/main/data/nlsy_kids.Rdata"))


```

```{r echo=TRUE}
# The PDF will show the code AND output here.
# Summarize the data here.



#looking at the amount of missing data to determine which ones are worth dropping.
colSums(is.na(nlsy_kids))/nrow(nlsy_kids)
#ppvt_3 is missing 80%+, dropping would result in most of the data being missing

nlsy_kids<-nlsy_kids %>% drop_na(lninc_0to3, momed, lnbw, 
                      comp_score_5to6 ,learndis)

view(nlsy_kids)

summary(nlsy_kids)


nlsy_kids_HS <- subset(nlsy_kids, head_start == 1)
nlsy_kids_NHS <- subset(nlsy_kids, head_start == 0)


#comparing the difference of mean and median of the two subsets of head_start

# Subset the dataset for head_start = 1
subset_head_start_1 <- subset(nlsy_kids, head_start == 1)

# Subset the dataset for head_start = 0
subset_head_start_0 <- subset(nlsy_kids, head_start == 0)

# Get the names of all variables in the dataset
all_variables <- names(nlsy_kids)

# Initialize vectors to store differences
mean_differences <- c()
median_differences <- c()

# Loop through variables
for (variable in all_variables) {
  # Skip non-numeric variables
  if (!is.numeric(nlsy_kids[[variable]])) {
    next
  }

  # Calculate mean and median for head_start = 1
  mean_head_start_1 <- mean(subset_head_start_1[[variable]])
  median_head_start_1 <- median(subset_head_start_1[[variable]])

  # Calculate mean and median for head_start = 0
  mean_head_start_0 <- mean(subset_head_start_0[[variable]])
  median_head_start_0 <- median(subset_head_start_0[[variable]])

  # Calculate differences
  mean_difference <- mean_head_start_1 - mean_head_start_0
  median_difference <- median_head_start_1 - median_head_start_0

  # Store differences in vectors
  mean_differences <- c(mean_differences, mean_difference)
  median_differences <- c(median_differences, median_difference)
}

# Create a data frame to display the results
results <- data.frame(
  Variable = all_variables,
  MeanDifference = mean_differences,
  MedianDifference = median_differences
)

# Print the results, head_start minus non_head_start
print(results)



```

\newpage

## Question 3

Code: Regression.

Verbal: Interpret.

Headstart has a -5.560 coefficient on test scores implying that by enrolling your child in a head start would result in a test score decrease by 5.6. With a t value of -4.888 the coefficient is statistically significant. It is reasonable to assume exogeneity, however it is likely there are omitted variables. We have to consider that the purpose of the headstart program is to help disadvantaged children with their education, so those enrolled in headstart may already have been on track to test lower scores down the line due to other environmental factors. This bias may be causing the coefficient to be much lower than it could be, since as it currently stands, the treatment seems worse than the symptom.


```{r echo=TRUE}
# All question 3 code here.

# Run OLS regression

ols_model <- feols(comp_score_5to6 ~ head_start, data = nlsy_kids)
summary(ols_model)

#boxplot(comp_score_5to6 ~ head_start, data = nlsy_kids, main = "Comparison of Test Scores by Head Start Participation")


```

\newpage

## Question 4

Code: Regression.

Verbal: Interpret.

The coefficient if still in the same order of magnitude and direction, but it has become larger in effect, -7.2604 compared to -5.560, showing that when accounting for the same family under the same mother, the effect has become stronger.

```{r echo=TRUE}
# All question 4 code here


# Create a new data frame with family-level means
family_means <- nlsy_kids %>%
  group_by(mom_id) %>%
  summarise(mean_head_start = mean(head_start),
            mean_test_scores = mean(comp_score_5to6))

# Run OLS regression
ols_model_family <- feols(mean_test_scores ~ mean_head_start, 
                          data = family_means)

# Summarize the model
summary(ols_model_family)


```

\newpage

## Question 5

Code: Regression.

Verbal: Interpret.

By using OLS with fixed effects of mother ID, the coefficient of head start becomes positive with a magnitude similar to the one found in question 4. Under the assumption that the Head start program should improve test scores, we would take this coefficient to be one that most likely reflects the effect of the program, as in previous analyses it has been returning coefficients that seem counter-intuitive to the intentions of the treatment.

```{r echo=TRUE}
# All question 5 code here

ols_model_fe <- feols(comp_score_5to6 ~ head_start | mom_id, 
                      data = nlsy_kids)
summary(ols_model_fe)



```

\newpage

## Question 6

Code: Regression.

Verbal: Interpret.

The coefficient of head start becomes negative again, but with a much smaller magnitude of 1.3095 demonstrating that when accounting for the ppvt test taken at age 3, head start has a much weaker predictive power on the age 5-6 test scores, and again has a negative effect as shown in questions 3 and 4. As the coefficient has swung to such a wide degree and has a standard error of 15 the robustness of the original regression may be called into question. When accounting separately for learning disability, robustness remains similar.

```{r echo=TRUE}
# All question 6 code here

ols_model_cov_ppvt <- feols(comp_score_5to6 ~ head_start + ppvt_3 | mom_id, 
                      data = nlsy_kids, vcov = "hetero")

ols_model_cov_learndis <- feols(comp_score_5to6 ~ head_start + learndis | mom_id, 
                      data = nlsy_kids, vcov = "hetero")

ols_model_cov_both <- feols(comp_score_5to6 ~ head_start + ppvt_3 + learndis | mom_id, 
                      data = nlsy_kids, vcov = "hetero")


summary(ols_model_cov_ppvt)
summary(ols_model_cov_learndis)
summary(ols_model_cov_both)


```

\newpage

## Question 7

Code: Regressions.

Verbal: Interpret.

The coefficients of head start on 7-10 is 0.166361 and 11-14 is 0.230717, both of which are lower than 5-6's 0.334273, showing that the effect seems to decrease with age, although not by much. the t-value for 7-10 is not high enough for statistical significance, so we can just focus on ages 11-14. The seemingly 33% drop in coefficient makes sense as the head start program was not a long lasting treatment and the students may forget what they have learned over time.

```{r echo=TRUE}
# All question 7 code here

# creating a function to standardize test scores across ages
standardize_scores <- function(scores) {
  return((scores - mean(scores, na.rm = TRUE)) / sd(scores, na.rm = TRUE))
}

# Standardize test scores for each age group
nlsy_kids$std_comp_score_5to6 <- standardize_scores(nlsy_kids$comp_score_5to6)
nlsy_kids$std_comp_score_7to10 <- standardize_scores(nlsy_kids$comp_score_7to10)
nlsy_kids$std_comp_score_11to14 <- standardize_scores(nlsy_kids$comp_score_11to14)

# Estimate fixed-effects models for each age group
fe_model_5to6 <- feols(std_comp_score_5to6 ~ head_start | mom_id, data = nlsy_kids, vcov = "iid")
fe_model_7to10 <- feols(std_comp_score_7to10 ~ head_start | mom_id, data = nlsy_kids, vcov = "iid")
fe_model_11to14 <- feols(std_comp_score_11to14 ~ head_start | mom_id, data = nlsy_kids, vcov = "iid")

# Print the summary of the models
summary(fe_model_5to6)
summary(fe_model_7to10)
summary(fe_model_11to14)


```

\newpage

## Question 8

Code: Regressions.

Verbal: Interpret.

When regressing on self-rated health, high school graduate and some college, some college is statistically insignificant. In regards self-rated health, the head start program has a negative effect. In regards to education, the effect on the probability to graduate high school is double that of attending college, 

```{r echo=TRUE}
# All question 8 code here

nlsy_kids <- nlsy_kids %>% drop_na(fphealth, hsgrad, somecoll)

# Estimate fixed-effects logit models for each outcome
fe_model_fphealth <- feols(fphealth ~ head_start | mom_id, data = nlsy_kids)
fe_model_hsgrad <- feols(hsgrad ~ head_start | mom_id, data = nlsy_kids)
fe_model_somecoll <- feols(somecoll ~ head_start | mom_id, data = nlsy_kids)

# Print the summary of the models
summary(fe_model_fphealth)
summary(fe_model_hsgrad)
summary(fe_model_somecoll)


```

\newpage

## Question 9

Code: Analysis of heterogeneity.

Verbal: Interpret.

```{r echo=TRUE}
# All question 9 code here

```

\newpage

## Question 9

Verbal: Policy implications.

